---
title: "Graph of Predictors for Weight Lifting"
author: "B Harold"
date: "2020-04-02"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Top Predictors Colored by Outcome

```{r echo=FALSE,results='hide',message=FALSE,warning=FALSE}
RNGversion("3.6.0")
library(caret)
library(plotly)
```


```{r cache=TRUE}
training <- read.csv("pml-training.csv", stringsAsFactors = FALSE)
```

```{r}
set.seed(1001)
minitrain <- createDataPartition(training$classe, p=0.1, list=FALSE)
tr2m = training[minitrain,]
tr2m$classe <- as.factor(tr2m$classe)
roll_belt <- tr2m$roll_belt
magnet_dumbbell_y <- tr2m$magnet_dumbbell_y
magnet_dumbbell_z <- tr2m$magnet_dumbbell_z
classe <- tr2m$classe
user_name <- tr2m$user_name
```

The dimensions here are the top three predictors in my previous study of this data, and the color indicates the outcome.  We see distinct clumping that is not directly related to outcome.

```{r}
plot_ly(
    x = ~roll_belt,
    y = ~magnet_dumbbell_y,
    z = ~magnet_dumbbell_z,
    color = ~classe,
    type = "scatter3d",
    mode = "markers",
    size = 4
)
```



## Top Predictors Colored by Subject

Here we color by the subject name instead, and this shows the source of this clumping.

```{r}
plot_ly(
    x = ~roll_belt,
    y = ~magnet_dumbbell_y,
    z = ~magnet_dumbbell_z,
    color = ~user_name,
    type = "scatter3d",
    mode = "markers",
    size = 4
)
```

## References

The data comes from this study.

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013. 

<http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har>

Plotted with plotly (<https://plotly.com>).
